The globular cluster system of the Auriga simulations

### 2019-08-19 AurigaGCS.tex
Edited abstract/introduction based on feedback WHT, RG, FM for #1. Textual remarks only, not yet feedback onto the content itself.

% Abstract of the paper
\begin{abstract}
We investigate whether the galaxy formation model used for the Auriga simulations
can produce a realistic globular cluster (GC) population at redshift zero. We
compare properties of the simulated star particles in the Auriga haloes with
catalogues of observations of the Milky Way (MW) and Andromeda (M31) globular
cluster populations available in the literature. We find that the Auriga 
simulations do produce sufficient mass at radii and metallicities that are 
typical for the MW globular cluster system (GCS), although we observe a varying 
mass-ratio for different galactocentric radius-metallicity 
($R_{\text{gal}}$-[Fe/H]) bins of simulated GC candidates
relative to the observed mass in the MW and M31 GC systems. This implies different 
values for the combined product of the bound cluster formation efficiency and 
the globular cluster disruption rate. Furthermore we test whether any of the 
Auriga galaxies has a metallicity and radial distribution that is consistent with 
the MW or M31 GCS. For any one of the Auriga haloes we reject the null hypothesis
that the observed and simulated metallicities of GCs (GC candidates) are drawn
from the same distribution at $\geq98.32$~\% confidence level, for the GCS of 
the Milky Way as well as that of the Andromeda galaxy. The same holds true for 
the distribution of $R_{\text{gal}}$ . Overall, the Auriga simulations produce 
old star particles with higher metallicities than the MW and M31 GCS and at 
larger radii. The formation efficiency would have to linearly decrease with 
increasing metallicity for the Auriga GC candidates to be consistent with the 
MW GCS, if cluster disruption is equally efficient regardless of metallicity.
In addition, the number of GCs relative to field stars would have to be smaller
for higher metallicity star particles at large radii than for those with lower 
metallicity at smaller radii, assuming GC disruption is more efficient in the
inner regions of the Galaxy.

\end{abstract}

@@

\section{Introduction}
Globular clusters (GC)s are old, bright, ubiquitous, and various properties of
GC systems show correlations with their host galaxies. GC systems are believed to
retain information about the galactic (gas) conditions at times of formation,
thus, could offer unique insight into the (chemodynamical) evolution of their 
parent galaxies, if the formation and evolution of GCs- and GC systems themselves 
is adequately understood. However, despite decades of research (the extensive 
literature on GCs is summarised in several books and review articles, e.g.~\citealt{
1991ARA&A..29..543H, Harris2001, 2004Natur.427...31W, 2006ARA&A..44..193B, 
2012A&ARv..20...50G, 2014CQGra..31x4006K, 2018RSPSA.47470616F}), consensus on an 
exhaustive picture of the formation of GCs is yet to be reached.

The literature offers a wealth of formation scenarios. \citet{1968ApJ...154..891P}
and \citet{1984ApJ...277..470P} argue that GCs form as the earliest bound
structures in the Universe (i.e. prior to formation of the main galaxy), 
noting that the Jeans length and mass shortly after recombination is consistent
with typical GC masses and sizes. \citet{2005MNRAS.364..367D} and \citet{
2009ApJ...706L.192B} argue that (the blue subpopulations of) GCs form in biased 
dark matter halos at high redshift. Other recent hypotheses of GC formation prior 
to collapse of the proto-galaxy include formation in colliding supershells 
\citep{2017Ap&SS.362..183R}, in supersonically induced gaseous objects
\citep{2019arXiv190408941C}, or in high-speed collisions of dark matter subhaloes
\citep{2019arXiv190508951M}. \citet{2017MNRAS.472.3120B} suggests
that (the blue subpopulation of) GCs could form in high density regions along
the cosmic filament before or during collapse.

Other models also date GCs formation during formation of the proto-galaxy itself,
for example as a result of thermal instabilities in hot gas-rich haloes 
(\citealt{1985ApJ...298...18F}, also see the discussions in \citealt{
1990ApJ...363..488K}). Alternative formation trigggers are also explored, such as 
(other causes of) shock compression, or cloud-cloud colissions \citep[e.g.][]{
1980glcl.conf..301G, 1992ApJ...400..265M, 1994ApJ...429..177H, 1995ApJ...442..618V,
1996ASPC...92..241L, 2001ApJ...560..592C}.

Yet another hypothesis is that star cluster formation is triggered by (major) 
gas-rich (spiral) galaxy mergers \citep{1987nngp.proc...18S, 1992ApJ...384...50A},
which is naturally expected within the framework of hierarchical assembly. One
testable prediction of this scenario is the formation of young clusters in
interacting and merging galaxies, which has been observed and are found to show 
remarkable similarities with globulars in the Milky Way \citep[e.g.][]{
1995AJ....109..960W, 1996AJ....112..416H, 1999AJ....118..752Z, 1999AJ....118.1551W}.
Moreover, modelling efforts of this framework yield GC (sub)populations consistent 
with various observables \citep[e.g.][]{2010ApJ...718.1266M, 2018MNRAS.480.2343C}, 
and the recent numerical simulation of an isolated dwarf-dwarf merger executed at 
very high resolution (baryonic mass $m_b \sim$~4M\Sun; softening $\epsilon = 0.1$) 
produce star clusters that could be globular progenitors \citep{
2019arXiv190509840L}.

As for the formation timeline, the scenarios (and flavours thereof) of \citet{
1985ApJ...298...18F} and \citet{1992ApJ...384...50A} are intertwined because
accretion and mergers continuously occurs trough out hierarchical build-up of
galaxies. Various (other) hierarchical formation channels thus combine different
aspects of the aforementioned paradigms, such as GC formation in (small) galactic
disks before they are accreted onto an assembling galaxy \citep[e.g.][]{
2000ApJ...533..869C, 2002ApJ...567..853C, 2002MNRAS.333..383B, 2003egcs.conf..224G}. 
We refer to \citet{2001astro.ph..8034G} for a scoreboard of the aforementioned 
paradigms compared to observations of the Milky Way globular cluster system.

An idea that has recently been studied in detail is the hypothesis that the physical 
processes responsible for the formation of young massive clusters \citep[YMCs, 
see][for a review]{2010ARA&A..48..431P}, as observable in detail in the local 
Universe in great
, also governs GC formation at high redshifts. In this framework, differences
between both classes of objects are caused by nearly a Hubble time of (dynamical)
evolution \citep[e.g.][]{1987degc.book.....S}. This picture is based on observed 
similarities between YMCs and GCs \citep[e.g.][]{1992AJ....103..691H,1999AJ....118.1551W}
and strengthened by observations of gravitationally lensed objects at high redshifts
($z = 2-6)$. These sources have properties reminiscent of (local) YMCs and may be
GC progenitors at times of formation \citep{2017MNRAS.467.4304V,2017ApJ...843L..21J}.
The modelling work by \citet{2011MNRAS.414.1339K,2012MNRAS.421.1927K,2015MNRAS.454.1658K} 
is now incorporated into cosmological zoom simulations and shows promising results
\citep{2018MNRAS.475.4309P,2019MNRAS.486.3134K}.

Simulation resolutions have reached the mass range populated by GCs, and the
gravitational force softening can be as low several parsec. A number of groups 
can thus incorporate formation of (globular) star clusters into their 
high-resolution hydrodynamical simulations. \citet{2016ApJ...831..204R} 
run parsec scale simulations of the high-redshift universe 
prior to reionization (the simulations stop at $z=9$), \citet{2017ApJ...834...69L} 
implement a new subgrid model for star (cluster) formation and run simulations 
that reach $z=3.3$, and the run of \citet{2017MNRAS.465.3622R} reaches $z=0.5$. 
\citet{2018MNRAS.474.4232K} find that mergers can push gas to high density that
quickly forms clustered stars that end up tightly bound by the end of the 
simulation. A somewhat different approach couples semi-analytical models to DM-only
simulations \citep{2010ApJ...718.1266M,2014ApJ...796...10L,2018MNRAS.480.2343C,
2019MNRAS.486..331C,2019arXiv190505199C}, as is done in the work by 
\citet{2019MNRAS.482.4528E}.

In this work we use state of the art simulations that produce realistic spiral 
galaxies at redshift zero for which several global properties are consistent with 
the observations. The simulations yield realistic Milky Way analogues, and real
galaxies host globular cluster systems. The question thus naturally arises whether 
the star formation histories of the simulations (also) give rise to a GC system 
similar to the Milky Way and/or Andromeda globular cluster systems. In particular, 
we use the Auriga suite of cosmological zoom simulations 
(\citealt{2017MNRAS.467..179G}, and described in Sec.~\ref{sec:auriga}) to
investigate whether the star formation model implemented produces metallicity, 
radial, and metallicity-radial distributions that are consistent with the MW 
and/or M31 GC systems, and whether the model produces enough stellar mass with
the right properties to allow for formation efficiencies lower than unity and
the expected (dynamical) mass loss over the cluster lifetime (not modelled 
explicitly in this study).

@@

The sub-grid includes primordial and metal-line cooling with self-shielding
corrections. Reionization is completed at redshift six by a time-varying
spatially uniform UV background \citep{2009ApJ...703.1416F, 2013MNRAS.436.3031V}
The interstellar medium is described by an equation of state for a two-phase medium
in pressure equilibrium \citep{2003MNRAS.339..289S} with stochastic star formation
in thermally unstable gas with a density threshold of $n = 0.13 \text{cm}^{-3}$,
and consecutive stellar evolution is accounted for. Stars provide feedback by
stellar winds \citep{2014MNRAS.437.1750M, 2017MNRAS.467..179G}, and further
enrich the ISM with metals from SNIa, SNII, and AGB stars \citep{2013MNRAS.436.3031V}.
The formation of black holes is modelled which results in feedback from active
galactic nuclei \citep{2005MNRAS.361..776S, 2014MNRAS.437.1750M, 2017MNRAS.467..179G}.
Finally, the simulations follow the evolution of a magnetic field of $10^{-14}$
(comoving)~G seeded at $z = 127$ \citep{2013MNRAS.432..176P, 2014ApJ...783L..20P}.
See G17 for further details of the numerical setup as well as the galaxy formation
model.

>>>>

% Abstract of the paper
\begin{abstract}
We investigate whether the galaxy formation model used for the Auriga simulations
can produce a realistic globular cluster (GC) population. We compare statistics 
of GC candidate star particles in the Auriga haloes with catalogues of the Milky 
Way (MW) and Andromeda (M31) globular cluster populations. We find that the Auriga 
simulations do produce sufficient stellar mass for GC candidates at radii and 
metallicities that are typical for the MW GC system (GCS), although we observe a 
varying mass-ratio for different galactocentric radius-metallicity 
($R_{\text{gal}}$-[Fe/H]) bins of simulated GC candidates relative to the 
observed mass in the MW and M31 GC systems.
This implies different values for the combined product of the bound cluster 
formation efficiency and the GC disruption rate. Furthermore we test whether any of the 
Auriga galaxies has a metallicity and radial distribution that is consistent with 
the MW or M31 GCS. For any one of the Auriga haloes we reject the null hypothesis
that the observed and simulated metallicities of GCs (GC candidates) are drawn
from the same distribution at $\geq98.32$~\% confidence level, for the GCS of 
the MW and M31. The same holds true for the distribution of $R_{\text{gal}}$.
Overall, the Auriga simulations produce old star particles with higher metallicities 
than the MW and M31 GCS and at larger radii. This implies that the bound cluster
formation efficiency would have to decrease with increasing metallicity for the 
Auriga GC candidates to be consistent with the MW GCS, if GC disruption is equally 
efficient regardless of metallicity. In addition, the number of GCs relative to
field stars would have to be smaller for higher metallicity star particles at
large radii than for those with lower metallicity at smaller radii, assuming GC 
disruption is more efficient in the inner regions of the Galaxy.
\end{abstract}

@@

\section{Introduction}
Globular clusters (GC)s are old, bright, and ubiquitous. Various properties of
GC systems show correlations with their host galaxies. GC systems are believed to
retain information about the galactic (gas) conditions at their times of formation.
Thus they could offer a unique insight into the (chemodynamical) evolution of their 
parent galaxies, if the formation and evolution of GCs- and GC systems themselves 
is adequately understood. However, despite decades of research (the extensive 
literature on GCs is summarised in several books and review articles, e.g.~\citealt{
1991ARA&A..29..543H, Harris2001, 2004Natur.427...31W, 2006ARA&A..44..193B, 
2012A&ARv..20...50G, 2014CQGra..31x4006K, 2018RSPSA.47470616F}), consensus on an 
exhaustive picture of the formation of GCs is yet to be reached.

The literature offers a wealth of formation scenarios. \citet{1968ApJ...154..891P}
and \citet{1984ApJ...277..470P} argue that GCs form as the earliest bound
structures in the Universe (i.e. prior to formation of the main galaxy), 
noting that the Jeans length and mass shortly after recombination is consistent
with typical GC masses and sizes. \citet{2005MNRAS.364..367D} and \citet{
2009ApJ...706L.192B} argue that (the blue subpopulations of) GCs form in radially
biased dark matter halos at high redshift. Other recent hypotheses of GC formation 
prior to collapse of the proto-galaxy include formation in colliding supershells 
\citep{2017Ap&SS.362..183R}, in supersonically induced gaseous objects
\citep{2019arXiv190408941C}, or in high-speed collisions of dark matter subhaloes
\citep{2019arXiv190508951M}. \citet{2017MNRAS.472.3120B} suggests
that (the blue subpopulation of) GCs could form in high density regions along
the cosmic filament before or during collapse.

Other models also date GCs formation during formation of the proto-galaxy itself,
for example as a result of thermal instabilities in hot gas-rich haloes 
(\citealt{1985ApJ...298...18F}, also see the discussions in \citealt{
1990ApJ...363..488K}). Alternative formation triggers are also explored, such as 
(other causes of) shock compression, or cloud-cloud colissions \citep[e.g.][]{
1980glcl.conf..301G, 1992ApJ...400..265M, 1994ApJ...429..177H, 1995ApJ...442..618V,
1996ASPC...92..241L, 2001ApJ...560..592C}.

Yet another hypothesis is that star cluster formation is triggered by major 
gas-rich (spiral) galaxy mergers \citep{1987nngp.proc...18S, 1992ApJ...384...50A},
which is naturally expected within the framework of hierarchical assembly. One
testable prediction of this scenario is the formation of young clusters in
interacting and merging galaxies, which has been observed and are found to show 
remarkable similarities with GCs in the Milky Way \citep[e.g.][]{
1995AJ....109..960W, 1996AJ....112..416H, 1999AJ....118..752Z, 1999AJ....118.1551W}.
Moreover, modelling efforts of this framework yield GC (sub)populations consistent 
with various observables \citep[e.g.][]{2010ApJ...718.1266M, 2018MNRAS.480.2343C}, 
and the recent numerical simulation of an isolated dwarf-dwarf merger executed at 
very high resolution (baryonic mass $m_b \sim$~4M\Sun; softening $\epsilon = 0.1$~pc) 
produces star clusters that could be globular progenitors \citep{
2019arXiv190509840L}.

As for the formation timeline, the scenarios (and flavours thereof) of \citet{
1985ApJ...298...18F} and \citet{1992ApJ...384...50A} are intertwined because
accretion and mergers continuously occurs during the hierarchical build-up of
galaxies. Various (other) hierarchical formation channels thus combine different
aspects of the aforementioned paradigms, such as GC formation in (small) galactic
disks before they are accreted onto an assembling galaxy \citep[e.g.][]{
2000ApJ...533..869C, 2002ApJ...567..853C, 2002MNRAS.333..383B, 2003egcs.conf..224G}. 
We refer to \citet{2001astro.ph..8034G} for a scoreboard of GC formation models 
compared to observations of the Milky Way globular cluster system.

An idea that has recently been studied extensively is the hypothesis that the formation
mechanisms of young massive clusters \citep[YMCs, see][for a review]{2010ARA&A..48..431P} 
and GCs are the same. The proximity of YMCs in the local Universe allows to study these 
processes with a level of detail that cannot be achieved for GCs at high redshifts. In 
this framework, differences between both classes of objects are caused by nearly a 
Hubble time of (dynamical) evolution \citep[e.g.][]{1987degc.book.....S}. This picture 
is based on observed similarities between YMCs and GCs
\citep[e.g.][]{1992AJ....103..691H,1999AJ....118.1551W} and strengthened by observations
of gravitationally lensed objects at high redshifts ($z = 2-6)$. These sources have 
properties reminiscent of (local) YMCs and may be GC progenitors at times of formation 
\citep{2017MNRAS.467.4304V,2017ApJ...843L..21J}. The modelling work by 
\citet{2011MNRAS.414.1339K,2012MNRAS.421.1927K,2015MNRAS.454.1658K} is now 
incorporated into cosmological zoom simulations and shows promising results
\citep{2018MNRAS.475.4309P,2019MNRAS.486.3134K}.

The (mass) resolution of cosmological zoom simulations has reached the mass range 
populated by GCs, and the gravitational force softening can be as low several 
parsec. A number of groups can thus incorporate formation of (globular) star 
clusters into their high-resolution hydrodynamical simulations. \citet{2016ApJ...831..204R} 
ran parsec scale simulations of the high-redshift universe 
prior to reionization (the simulations stop at $z=9$), \citet{2017ApJ...834...69L} 
implement a new subgrid model for star (cluster) formation and run simulations 
that reach $z=3.3$, and the run of \citet{2017MNRAS.465.3622R} reaches $z=0.5$. 
\citet{2018MNRAS.474.4232K} find that mergers can push gas to high density that
quickly forms clustered stars that end up tightly bound by the end of the 
simulation. A somewhat different approach couples semi-analytical models to DM-only
simulations \citep{2010ApJ...718.1266M,2014ApJ...796...10L,2018MNRAS.480.2343C,
2019MNRAS.486..331C,2019arXiv190505199C}, as is done in the work by 
\citet{2019MNRAS.482.4528E}.

In this work we use state of the art simulations that produce realistic spiral 
galaxies (Milky Way analogues) at redshift zero for which several global properties 
are consistent with the observations. The question thus naturally arises whether 
the star formation histories of the simulations give rise to a GC system 
similar to the Milky Way and/or Andromeda globular cluster systems. In particular, 
we use the Auriga suite of cosmological zoom simulations 
(\citealt{2017MNRAS.467..179G}, and described in Sec.~\ref{sec:auriga}) to
investigate whether the star formation model implemented produces metallicity, 
radial, and metallicity-radial distributions that are consistent with the MW 
and/or M31 GC systems, and whether the model produces enough stellar mass with
the right properties to allow for formation efficiencies lower than unity and
the expected (dynamical) mass loss over the cluster lifetime (not modelled 
explicitly in this study).

@@

The sub-grid includes primordial and metal-line cooling with self-shielding
corrections. Reionization is completed at redshift six by a time-varying
spatially uniform UV background \citep{2009ApJ...703.1416F, 2013MNRAS.436.3031V}.
The interstellar medium is described by an equation of state for a two-phase medium
in pressure equilibrium \citep{2003MNRAS.339..289S} with stochastic star formation
in thermally unstable gas with a density threshold of $n = 0.13 \text{cm}^{-3}$,
and consecutive stellar evolution is accounted for. Stars provide feedback by
stellar winds \citep{2014MNRAS.437.1750M, 2017MNRAS.467..179G}, and further
enrich the ISM with metals from SNIa, SNII, and AGB stars \citep{2013MNRAS.436.3031V}.
The formation of black holes is modelled which results in feedback from active
galactic nuclei \citep{2005MNRAS.361..776S, 2014MNRAS.437.1750M, 2017MNRAS.467..179G}.
Finally, the simulations follow the evolution of a magnetic field of $10^{-14}$
(comoving)~G seeded at $z = 127$ \citep{2013MNRAS.432..176P, 2014ApJ...783L..20P}.
See G17 for further details of the numerical setup as well as the galaxy formation
model.
